$\pi$ with leftovers: \\ a mechanisation in Agda

### 2020-05-04 paper/paper.tex
Get rid of rearrangement figure

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\todo{highlight that we can talk about a resource even if it has been exhausted (e.g. we can send 0 multiplicities of a channel over another channel}
\todo{check 'z' are replaced with 's' in formaliz*, mechaniz*; check modeling is modelling}
\todo{rename section and subsections in reference to \S (not sure why the commands above are not doing it so check later)}
We live in a concurrent world where any given state is often interactively computed by a myriad of parties --- people, machines, processors, services, networks.
As humans, we aim to model, predict, and build such interactive systems; as mathematicians, abstraction is our tool of choice to do so.

The \picalc{} \cite{MilnerPW92,Milner99} is the most successful computational model for communication and concurrency.
It abstracts over the details of concurrent processing and boils the interactions down to the sending and receiving of data over communication channels.
Notably, it features channel mobility: channels themselves are first order values, and can therefore be transmitted.
The \picalc{} has been a foundation for the design and implementation of programming languages for concurrency, such as Pict \cite{Pierce} and (more recently) Go \cite{Golang}.
At the state-of-the-art, the \picalc{} features a wide plethora of types \cite{K07}: basic types, linear types, types for liveness properties (such as deadlock freedom, livelock freedom or termination), and most notably, session types. This makes the \picalc{} a fully-fledged tool for modeling and verification of concurrent and distributed systems.

@@

  The base case $\constr{0}$ splits the usage annotation $x$ of type $\Usage_{idx}$ into $y$ and $z$, which are the leftovers.
  This splitting $\opsquared{x}{y}{z}$ is as per the usage algebra provided by the developer for the index $idx$.
  We use the functionality of the monoidal relation to alleviate the user from this proof burden.
  The inductive case $\suc$ appends the type $t'$ to the typing context, and the usage annotation $x'$ to both the input and output usage contexts.

@@

\todo{maybe only focus on the first case, internal and external and say the rest is by induction blah bla and details can be found in appenix/or github}
\begin{proof}
  \hfill{}\\
  \begin{itemize}
    \item
    In the $\constr{comm}$ case we apply framing (\autoref{thm:framing}) to rearrange the assumptions in \autoref{fig:subject-reduction:a} into the proofs in \autoref{fig:subject-reduction:b}.
    We then apply substitution (\autoref{thm:substitution}) and strengthening (\autoref{thm:strengthening}).

    \begin{figure}[h]
      \begin{subfigure}{.5\textwidth}
        \begin{alignat*}{2}
          \containsusage{&\Gamma'}{i}{\lio &&}{\Gamma} \\
          \contains{\gamma}{&\Gamma'}{i}{\channel{t}{m}}{\li &&}{\Psi} \\
          \types{\gamma \comma t}{&\Psi \comma m}{P &&}{\Delta \comma \lz} \\
          \contains{\gamma}{&\Delta }{i}{\channel{t}{m}}{\lo &&}{\Theta} \\
          \contains{\gamma}{&\Theta}{j}{t}{m &&}{\Xi}
        \end{alignat*}
        \caption{Initial assumptions.}
        \label{fig:subject-reduction:a}
      \end{subfigure}
      % Do not leave empty space
      \begin{subfigure}{.5\textwidth}
        \begin{alignat*}{2}
          \types{\gamma \comma t}{& \Gamma \comma m}{P &&}{\Theta \comma \lz} \\
          \contains{\gamma}{& \Theta}{j}{t}{m &&}{\Xi}
        \end{alignat*}
        \caption{Shift witness for $\lo$ and merge with the one for $\li$ into $\lio$.}
        \label{fig:subject-reduction:b}
      \end{subfigure}
    \end{figure}

@@

As future work we intend to...
Adding product types and sum types.
Proving that progress preserves the balancing of channels.
Erika Kreuer has suggested modeling the usage algebra as a state machine.
Adding bidirectionality would provide us with a decision procedure for type checking for a given context of algebras.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\todo{highlight that we can talk about a resource even if it has been exhausted (e.g. we can send 0 multiplicities of a channel over another channel}
\todo{check 'z' are replaced with 's' in formaliz*, mechaniz*; check modelling is modelling}
\todo{rename section and subsections in reference to \S (not sure why the commands above are not doing it so check later)}
We live in a concurrent world where any given state is often interactively computed by a myriad of parties --- people, machines, processors, services, networks.
As humans, we aim to model, predict, and build such interactive systems; as mathematicians, abstraction is our tool of choice to do so.

The \picalc{} \cite{MilnerPW92,Milner99} is the most successful computational model for communication and concurrency.
It abstracts over the details of concurrent processing and boils the interactions down to the sending and receiving of data over communication channels.
Notably, it features channel mobility: channels themselves are first order values, and can therefore be transmitted.
The \picalc{} has been a foundation for the design and implementation of programming languages for concurrency, such as Pict \cite{Pierce} and (more recently) Go \cite{Golang}.
At the state-of-the-art, the \picalc{} features a wide plethora of types \cite{K07}: basic types, linear types, types for liveness properties (such as deadlock freedom, livelock freedom or termination), and most notably, session types. This makes the \picalc{} a fully-fledged tool for modelling and verification of concurrent and distributed systems.

@@

  The base case $\constr{0}$ splits the usage annotation $x$ of type $\Usage_{idx}$ into $y$ and $z$, which are the leftovers.
  This splitting $\opsquared{x}{y}{z}$ is as per the usage algebra provided by the developer for the index $idx$.
  We use the \emph{functionality} of the monoidal relation to alleviate the user from the proof burden $\opsquared{x}{y}{z}$.
  The inductive case $\suc$ appends the type $t'$ to the typing context, and the usage annotation $x'$ to both the input and output usage contexts.

@@

\todo{maybe only focus on the first case, internal and external and say the rest is by induction blah bla and details can be found in appenix/or github}
\begin{proof}
  \hfill{}\\
  \begin{itemize}
    \item
    In the $\constr{comm}$ case we apply framing (\autoref{thm:framing}) to rearrange the assumptions and then apply substitution (\autoref{thm:substitution}) and strengthening (\autoref{thm:strengthening}).

@@

As future work we intend to...
Adding product types and sum types.
Proving that progress preserves the balancing of channels.
Erika Kreuer has suggested modelling the usage algebra as a state machine.
Adding bidirectionality would provide us with a decision procedure for type checking for a given context of algebras.
